---
title: "methylation at siRNA loci"
author: "Chenxin Li"
date: "12/29/2020"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(emmeans) 
library(stringr)
library(RColorBrewer)
library(svglite) 
```

# load data 

read_delim("C:/Users/cxli9/Desktop/Li/PBI/PBI299 (Sundar Lab)/smRNA project/CLVS13/drm2_ATCGmaps/wt3_em_S47_L005_R1_001.fastq.gz.trimmed.fastq.gz.filtered.fastq.gz.sorted.bam.CGmap.gz.CHH.zygote-egg_intersect_loci.mtr", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE) 
    
    
```{r message=F, warning = F}
em_meth_list <- list.files(path = "C:/Users/cxli9/Desktop/Li/PBI/PBI299 (Sundar Lab)/smRNA project/CLVS13/drm2_ATCGmaps/", 
                        pattern = "*.mtr", full.names = T) 

em_meth_data <- sapply(em_meth_list, read_delim, simplify = FALSE, 
                       delim = "\t", col_names = F, escape_double = F, trim_ws = T) %>% 
  bind_rows(.id = "id")

head(em_meth_data) 
```

# keying data 

each locus must have 3 or more reads to be considered in the analyses 
```{r}
em_meth_data_nice <- em_meth_data %>% 
  filter(X7 >= 3) %>% 
   mutate(sample_ID = case_when(
    str_detect(id, "52_38") ~ "drm2_38", 
    str_detect(id, "52_39") ~ "drm2_39", 
    str_detect(id, "52_40") ~ "drm2_40", 
    str_detect(id, "wt1_em") ~ "WT1", 
    str_detect(id, "wt2_em_comb") ~ "WT2", 
    str_detect(id, "wt3_em") ~ "WT3"
  )) %>% 
  filter(is.na(sample_ID) == F) %>% 
  mutate(locus = case_when(
    str_detect(id, "egg_loci") ~ "egg siRNA loci", #1
    str_detect(id, "seedling_loci") ~ "seedling siRNA loci", #2
    str_detect(id, "embryo_loci") ~ "embryo siRNA loci", #3 
    str_detect(id, "zygote-egg_complement_loci.") ~ "zygote NOT egg", #4
    str_detect(id, "zygote-egg_intersect") ~ "zygote/egg intersect",  #5
    str_detect(id, "sperm_loci") &
      str_detect(id, "endo") == F ~ "sperm siRNA loci", #6 
    str_detect(id, "egg-zygote_complement_loci.") ~ "egg NOT zygote", #7 
    str_detect(id, "endosperm_loci.") ~ "endosperm siRNA loci", #8
    str_detect(id, "zygote_loci.") ~ "zygote siRNA loci", #9
    str_detect(id, "egg_siren") ~ "egg siren loci" #10 
  )) %>% 
  filter(is.na(locus) == F) %>% 
  mutate(context = case_when(
    str_detect(id, "gz.CG.") ~ "mCG", 
    str_detect(id, "gz.CHG.") ~ "mCHG", 
    str_detect(id, "gz.CHH.") ~ "mCHH"
  )) %>% 
  mutate(genotype = case_when(
    str_detect(sample_ID, "drm2") ~ "drm2",
    str_detect(sample_ID, "WT") ~ "WT"
  )) 

head(em_meth_data_nice)
```
 

# stats

```{r}
em_meth_data_nice %>% 
  group_by(locus, context, sample_ID, genotype) %>% 
  count
```

## for each locus, average bioreps to the level of genotype
```{r}
em_meth_data_s <- em_meth_data_nice %>% 
 filter(is.na(X4) == F) %>% 
 group_by(X1, X2, X3, locus, context, genotype) %>% 
 summarise(meth = mean(X4)) %>% 
 ungroup()

head(em_meth_data_s)
```
## summary stats - quantiles 
```{r}
em_meth_data_ss <- em_meth_data_s %>% 
  filter(is.na(meth) == F) %>% 
  group_by(locus, context, genotype) %>% 
  summarise(
    NN = n(),
    median = median(meth),
    Q3 = quantile(meth, 0.75),
    Q1 = quantile(meth, 0.25),
    UL = quantile(meth, 0.975),
    LL = quantile(meth, 0.025)
  ) %>% 
  ungroup()

em_meth_data_ss
```
```{r}
em_meth_data_ss %>% 
  filter(context == "mCHH") %>% 
  filter(genotype == "drm2")
```
* Zygote NOT egg - egg NOT zygote = +30% 
* zygote - egg = +22% 

## linear models
```{r}
logit <- function(p){log(
  p / (1-p)
)}

logistic <- function(x){
  1/(1 + exp(-x))
}
```

```{r}
model_em_meth <- glm(meth ~ context * genotype * locus, data = em_meth_data_s %>% 
                       filter(locus != "endosperm siRNA loci") %>% 
                       filter(locus != "egg siren loci"), 
                     family = quasibinomial(link = "logit")) 

anova(model_em_meth, test = "LRT")
```

```{r}
est_em_meth1 <- emmeans(model_em_meth, pairwise ~ genotype | locus | context) 
est_em_meth2 <- emmeans(model_em_meth, pairwise ~ locus | genotype | context) 
```

```{r}
em_meth_results <- multcomp::cld(est_em_meth2$emmeans, Letters = letters) %>% 
  as.data.frame() %>% 
  mutate(grouping = str_remove_all(.group, " ")) 

em_meth_results %>% 
  filter(context == "mCHH") %>% 
  select(locus, genotype, context, grouping)
```

```{r}
est_em_meth2$contrasts %>% 
  as.data.frame() %>% 
  filter(genotype == "WT") %>% 
  filter(context == "mCHH") %>% 
  select(contrast, p.value)
```

* egg NOT zygote - zygote NOT egg: P = 0
* egg siRNA loci - zygote siRNA loci: P = 0 

# plot
## CHH 
```{r}
em_meth_data_ss %>% 
  filter(context == "mCHH") %>%
  filter(locus != "endosperm siRNA loci") %>% 
  filter(locus != "egg siren loci") %>% 
  inner_join(em_meth_results, by = c("locus", "context", "genotype")) %>% 
  mutate(letter_pos = case_when(
    genotype == "WT" ~ median,
    T ~ Q3
  )) %>% 
  mutate(vjust = case_when(
    genotype == "WT" ~ 0.2,
    T ~ -0.3
  )) %>%
  mutate(locus2 = case_when(
    str_detect(locus, "embryo siRNA loci") ~ "*embryo siRNA loci",
    T ~ locus
  )) %>% 
  mutate(locus = reorder(locus2, median)) %>% 
  ggplot(aes(x = locus, y = Q3)) +
  facet_grid(genotype ~., scales = "free", switch = "y") + 
  geom_crossbar(aes(ymin = Q1, ymax = Q3, y = median), width = 0.7, alpha = 0.8) +
  geom_segment(aes(x = locus, xend = locus,  y = Q3, yend = UL)) +
  geom_segment(aes(x = locus, xend = locus,  y = Q1, yend = LL)) +
  geom_errorbar(aes(ymin = UL, ymax = UL), width = 0.3, size = 1) +
  geom_errorbar(aes(ymin = LL, ymax = LL), width = 0.3, size = 1) +
  geom_text(aes(label = grouping, y = letter_pos, vjust = vjust),
   hjust = -0.5, size = 4, fontface = "bold") +
  labs(x = NULL, 
       y = "mCHH in WT or drm2 embryo") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, face = "bold", colour = "black"),
    axis.text = element_text(colour = "black"), 
    axis.text.y = element_text(hjust = 0.5),
    strip.placement = "outside",
    strip.text.y = element_text(hjust = 0.5, angle = 0, face = "bold"), 
    panel.spacing = unit(1, "lines"),
    axis.line = element_line(size = 1),
    legend.position = "none"
  ) +
  coord_flip()  

ggsave("em_meth.svg", height = 6, width = 6)
ggsave("em_meth.png", height = 6, width = 6)
```
 

## CHG 
```{r}
em_meth_data_ss %>% 
  filter(context == "mCHG") %>%
  filter(locus != "endosperm siRNA loci") %>% 
  filter(locus != "egg siren loci") %>% 
  inner_join(em_meth_results, by = c("locus", "context", "genotype")) %>% 
  mutate(letter_pos = case_when(
    genotype == "WT" ~ median,
    T ~ Q3
  )) %>% 
  mutate(vjust = case_when(
    genotype == "WT" ~ 0.2,
    T ~ -0.3
  )) %>%
  mutate(locus2 = case_when(
    str_detect(locus, "embryo siRNA loci") ~ "*embryo siRNA loci",
    T ~ locus
  )) %>% 
  mutate(locus = factor(locus2, levels = c(
    "sperm siRNA loci",
    "egg NOT zygote",
    "egg siRNA loci",
    "*embryo siRNA loci",
    "zygote/egg intersect",
    "zygote siRNA loci",
    "zygote NOT egg",
    "seedling siRNA loci"
  ))) %>% 
  ggplot(aes(x = locus, y = Q3)) +
  facet_grid(genotype ~., scales = "free", switch = "y") + 
  geom_crossbar(aes(ymin = Q1, ymax = Q3, y = median), width = 0.7, alpha = 0.8) +
  geom_segment(aes(x = locus, xend = locus,  y = Q3, yend = UL)) +
  geom_segment(aes(x = locus, xend = locus,  y = Q1, yend = LL)) +
  geom_errorbar(aes(ymin = UL, ymax = UL), width = 0.3, size = 1) +
  geom_errorbar(aes(ymin = LL, ymax = LL), width = 0.3, size = 1) +
  geom_text(aes(label = grouping, y = letter_pos, vjust = vjust),
   hjust = -0.5, size = 4, fontface = "bold") + 
  labs(x = NULL, 
       y = "mCHG in WT or drm2 embryo") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, face = "bold", colour = "black"),
    axis.text = element_text(colour = "black"), 
    strip.placement = "outside",
    strip.text.y = element_text(hjust = 0.5, angle = 0, face = "bold"), 
    panel.spacing = unit(1, "lines"),
    axis.line = element_line(size = 1),
    legend.position = "none"
  ) +
  coord_flip()  

ggsave("em_meth_CHG.svg", height = 6, width = 6)
ggsave("em_meth_CHG.png", height = 6, width = 6)
```

## GG
```{r}
em_meth_data_ss %>% 
  filter(context == "mCG") %>%
  filter(locus != "endosperm siRNA loci") %>% 
  filter(locus != "egg siren loci") %>% 
  inner_join(em_meth_results, by = c("locus", "context", "genotype")) %>% 
  mutate(letter_pos = case_when(
    genotype == "WT" ~ median,
    T ~ Q3
  )) %>% 
  mutate(vjust = case_when(
    genotype == "WT" ~ 0.2,
    T ~ -0.3
  )) %>%
  mutate(locus2 = case_when(
    str_detect(locus, "embryo siRNA loci") ~ "*embryo siRNA loci",
    T ~ locus
  )) %>% 
  mutate(locus = factor(locus2, levels = c(
    "sperm siRNA loci",
    "egg NOT zygote",
    "egg siRNA loci",
    "*embryo siRNA loci",
    "zygote/egg intersect",
    "zygote siRNA loci",
    "zygote NOT egg",
    "seedling siRNA loci"
  ))) %>% 
  ggplot(aes(x = locus, y = Q3)) +
  facet_grid(genotype ~., scales = "free", switch = "y") + 
  geom_crossbar(aes(ymin = Q1, ymax = Q3, y = median), width = 0.7, alpha = 0.8) +
  geom_segment(aes(x = locus, xend = locus,  y = Q3, yend = UL)) +
  geom_segment(aes(x = locus, xend = locus,  y = Q1, yend = LL)) +
  geom_errorbar(aes(ymin = UL, ymax = UL), width = 0.3, size = 1) +
  geom_errorbar(aes(ymin = LL, ymax = LL), width = 0.3, size = 1) +
  geom_text(aes(label = grouping, y = Q1),
   hjust = 1.5, size = 4, fontface = "bold", vjust = -0.3) + 
  labs(x = NULL, 
       y = "mCG in WT or drm2 embryo") +
  theme_minimal() +
  theme(
    text = element_text(size = 14, face = "bold", colour = "black"),
    axis.text = element_text(colour = "black"), 
    strip.placement = "outside",
    strip.text.y = element_text(hjust = 0.5, angle = 0, face = "bold"), 
    panel.spacing = unit(1, "lines"),
    axis.line = element_line(size = 1),
    legend.position = "none"
  ) +
  coord_flip()  

ggsave("em_meth_CG.svg", height = 6, width = 6)
ggsave("em_meth_CG.png", height = 6, width = 6)
```


 

